An interdisciplinary examination of stress and injury occurrence in athletes

### 2020-07-29 apa_style.Rmd
progress

floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
csl: frontiers.csl
always_allow_html: true
---

@@

```{r include=FALSE}
options(scipen=999, digits = 2)
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
# source("r-files/functions.R")
# source("r-files/methods_code.R")
# source("r-files/results_code.R")
load("all_data.RData")
pacman::p_load(tidyverse, knitr, kableExtra, qwraps2, brms, broom.mixed, bnlearn, tidyselect, Rgraphviz)
library(readxl)
library(janitor)
library(irr)
library(lavaan)
library(caret)
library(parallel)
library(zoo)
library(sjmisc)

@@

\section{Introduction}

@@

@@

\section{Methods}

@@

\subsection{Participants}

A total of 351 athletes (male: _n_ = 231, female: _n_ = 120) were initially recruited for the study from a British University and local sports clubs (Table \ref{tab:table1}). Participants had an average age of `r mean_sd(parchar$age, digits = 1)` years and represented a range of team (football, rugby, netball, cricket, lacrosse, basketball and field hockey) and individual sports (athletics, tennis, weightlifting, gymnastics, judo, swimming and golf. 
Participants self-rated competitive level ranged from recreational to international standard. 
They were engaged in training for their respective sports for at least five hours per week.
A total of `r n_perc(parchar$pic == "1")` participants had sustained an injury in the 12 months prior to the start of the study (male: _n_ = 114 [49%], female; _n_ = 48 [40%]).
All participants were injury free (no modifications to their usual training routine due to a sport related medical problem for a minimum of four weeks) at the start of the study. 
Ethical approval was obtained from the University ethics committee prior to the start of the study and all participants provided informed consent.

@@

@@

@@

@@

@@

BNs have recently been used in the sport psychology literature [@Olmedilla2018; @Fuster-Parra2017] and offer several benefits over traditional statistical analysis. 
For example, predictions can be made about any variable in the network, rather than there being a distinction between dependent and independent variables in the data, such as in linear regression models that are often used within the sport psychology literature [@Olmedilla2018; @Bittencourt2016].
Furthermore, the structure of a network can be obtained from both empirical data _and_ prior knowledge about the area of study; the latter being particularly useful when there are a large number of variables in the network, or only a small number of observations are available in the data [@Xiao-xuan2007].
In such instances, a purely data driven approach to learning the network would be time-consuming due to the large parameter space, and inefficiency at identifying an approximation of the true network structure. 
Prior knowledge about dependencies between variables can therefore be included in the network structure, while still allowing a data driven approach for unknown dependencies, to improve the overall computation of the network structure [@Heckerman1995; @Xu2015].
The following sections detail the steps taken in the current study to firstly prepare the data for each network, and then obtain the structure of each network that was used for further inference.

@@

Of the 351 participants that were initially recruited for the study, 94 only completed the first time point, and therefore had to be removed from the study as no injury information was obtained for these participants following the first time point. 
To prepare the data for the BN, missing values in the dataset were first imputed. 
Out of the 650 total measurements across all time points in the current study, there were `r n_perc(missingtot$stiffness == 0)` missing muscle stiffness measurements and `r n_perc(missingtot$rmssd == 0)` missing heart rate recordings. The missing data were due to technical faults in the data collection equipment and were considered to be missing completely at random. 
A missing rate of 15-20% has been reported to be common in psychological studies, and several techniques are available to handle missing values [@Enders2003; @Lang2014].
In the current study, the _caret_ package [@Kuhn2008] was used to impute the missing values. 
A bagged tree model using all of the non-missing data was first generated and then used to predict each missing value in the dataset. 
The bagged tree method is a reliable and accurate method for imputing missing values in data and is superior to other commonly used methods such a median imputation [@Kuhn2008].

A median split technique was used to discretise the data used in the network into "Low" and "High" levels.
All variables apart from negative and total life events were approximately normally distributed (based on visual inspection, see supplementary material Sx) and required no further transformation prior to the median split. 
For the LESCA questionnaire data, a cumulative total of the current, and previous time points was calculated at each time point to account for the potential continuing effect of the life events experienced by athletes over time.
Given the limited support for a relationship between positive life events and injury [@Williams2007], only negative and total life events were included in the network.
Cumulative negative, and cumulative total life event scores at each time point were first log scaled so distributions were approximately normal, and then binarised using the median at each time point (nlelg and tlelg respectively).
In addition to the log scaled cumulative values, an untransformed NLE score from the first time point was included as an additional variable based on previous literature that indicates this variable should have a strong relationship with injury outcome [@Ivarsson2017].

@@

To obtain the network structure, several steps were taken to ensure that both a theoretically realistic network, and a network that was an appropriate fit to the collected data, was used for inference. 
Prior knowledge about the network structure was included by providing a list of arcs that are always _restricted_ from being in the network (blacklist), and a list of arcs that are always _included_ in the network (whitelist). 
Additionally, there are several scoring functions such as Bayesian Information Criteria (BIC) and Bayesian Dirichlet equivalent uniform (BDeu) that can be used to compare network structures with certain nodes and arcs included or excluded [@Scutari2014]. 
To account for the repeated measures design employed in this study and to maximise the use of the data, pairs of complete cases (e.g., participants who completed T1 + T2, and T2 + T3) were used in a two-time Bayesian network (2TBN) structure [@Murphy2002].
In the 2TBN, variables measured T2 could depend on variables measured at T1 (e.g., T1 $\rightarrow$ T2) and variables measured at T3 could depend on variables measured at T2 (e.g., T2 $\rightarrow$ T3).
However, arcs were blacklisted between T2 $\rightarrow$ T1 and T3 $\rightarrow$ T2 to preserve the order in which data was collected. 
Variables were separated into two groups; "explanatory", for variables that did not change during the study (e.g., gender), or "independent", for variables that were measured at each time point and could vary during the study.
Independent variable names were suffixed with _1 for time point T, and _2 for time point T+1 (e.g., T1_1 $\rightarrow$ T2_2 and T2_1 $\rightarrow$ T3_2).
Formatting the data in this way meant participants who completed T1 and T2, but did not complete T3, could still be included in the analysis. 
Table \ref{tab:table2} provides an example of the formatted data and demonstrates that participants 1 and 3 have complete data, and therefore have two rows of data each representing variables from T1 $\rightarrow$ T2 and T2 $\rightarrow$ T3, respectively. 
Participant 2 did not complete the final data collection at T3 and therefore only has one row of data representing the variables collected at T1 and T2.
In addition to the blacklisted arcs between T2 $\rightarrow$ T1 and T3 $\rightarrow$ T2, the direction of arcs was restricted between independent variables and explanatory variables (e.g., independent $\rightarrow$ explanatory); however, arcs were not restricted between explanatory $\rightarrow$ independent variables.
Finally, arc direction was restricted between specific nodes within the explanatory variables. 
Arcs from clevel $\rightarrow$ gender, nlebase $\rightarrow$ gender and nlebase $\rightarrow$ ind_team were included in the blacklist, as arcs in these directions did not make logical sense. 
All subsequent models used the same blacklist.

@@

The subscales representing the BAS (RR, RI, GDP and I) showed limited connection to other variables in the network. 
Therefore, several models were run with each scale individually to find the scale that resulted in the highest BIC value (values are shown relative to the highest value).
RI provided the highest BIC value, compared to RR (`r rstmodel[2,2]- rstmodel[1,2]`), GDP (`r rstmodel[4,2]-  rstmodel[1,2]`) and I (`r rstmodel[5,2] - rstmodel[1,2]`).
Including all the variables (RR, RI, GDP and I) resulted in a significantly lower score `r rstmodel[6,2] - rstmodel[1,2]`) indicating that including all the variables was not beneficial to the model structure and did not offset the cost of the additional parameters. Therefore, only RI was included in the final structure.

Finally, both total score and asymmetry for balance were included in the initial network. However, visual inspection of the network revealed no arcs between bal_asym_1 or bal_asym_2 and any other node in the network.
Therefore, balance asymmetry was removed from the final network structure.
To summarise, Table \ref{tab:table3} reports the variables that were included in the final network structure.

```{r table3, results='asis'}
varcutoffs %>% 
  filter(!var %in% c("sdnn", "RR", "I", "GDP", "tlelg", "tlelg_1", "tlelg_2", "tlelg_3", "bal_asym")
) %>% 
  mutate(var = ifelse(var == "rmssd", "HRV", var)) %>% 
  kable("latex",
       align = c("llcc"),
       caption = "Variables included in the final Bayesian network structure.",
       col.names = c("Variable", "Definition", "State 1", "State 2")) %>%
  kable_styling(latex_options = "HOLD_position") %>% 
  column_spec(2, width = c("5.2cm")) 
```

@@

During the study, 26% of participants reported at least one injury with an average severity of `r mean_sd(datanet$days_missed, digits = 0)`, days (range = `r min(datanet$days_missed[datanet$days_missed > 1])` - `r max(datanet$days_missed)` days).
Both male and female participants reported a greater number of acute compared to chronic injuries (male, acute = `r inj_mechanism[1,3]`, chronic = `r inj_mechanism[2,3]`; female, acute = `r inj_mechanism[1,2]` chronic = `r inj_mechanism[2,2]`), and non-contact injuries were more common than contact injuries (male, non-contact = `r inj_mechanism[4,3]`, contact = `r inj_mechanism[2,3]`; female, non-contact = `r inj_mechanism[4,2]` contact = `r inj_mechanism[3,2]`).
Table \ref{tab:table4} shows the number and percentage of injury types sustained by both male and female participants.

@@

y <- pmap_dfr(.l = list(from = rep(from, 2),
     to = rep(to, 2),
     state = rep(c("High", "Low"), each = length(from))),
         ~newprobtable(tallc, vars = ..1, outcome = ..2, state = ..3, avg30)) %>%
  gather(var, state, -prob, -outcome) %>%
  drop_na(state) %>%
  mutate(prob = format(round(prob, digits = 2), nsmall = 2))
```

@@

cbind(inj_summary[c(1:3), c(2:4)], inj_summary[c(4:6), c(3:4)]) %>%
  kable(
    .,
    caption = "The number and percentage (\\%) of types of injuries sustained by male and female participants.",
    col.names = c("", "Lower body", "Upper body", "Lower body", "Upper body"),
    escape = F
  ) %>%
  add_header_above(c(" " = 1, "Female" = 2, "Male" = 2), bold = T) %>%
  footnote(general = "Other included bone, skin and brain injuries.") %>%
  kable_styling(latex_options = "HOLD_position")  %>%
  row_spec(0,bold=TRUE)

@@

The first network structure obtained from the data is shown in Fig \ref{fig:fig6}. 
Several of the explanatory variables showed strong connections with independent variables in the network.
The arc from nlebase $\rightarrow$ RI_1 had a strength of 0.84, and the probability of RI_1 being in the "High" state increased from `r x[1,1]` to `r x[2,1]` when nlebase increased from "Low" to "High".
The ind_team node had strong arcs to hours (0.90) and nlebase (0.84).
Individual athletes were more likely to have "High" hours per week (`r x[5,1]`) compared to team-based athletes (`r x[6,1]`).
Individual athletes were also more likely to have "High" negative life events in the 12 months preceding the start of the study compared to team based athletes (individual athletes = `r x[7,1]`, team-based athletes = `r x[8,1]`).
The arcs from gender $\rightarrow$ stiffness_1 and gender $\rightarrow$ stiffness_2 were 0.76, and 0.65 respectively, with males more likely to have "High" stiffness compared to females (males = `r x[14,1]`, females = `r x[13,1]`).
The arc from pi $\rightarrow$ stiffness_1 was 0.55 with athletes who reported an injury in the preceding 12 months more likely to have "High" (`r x[26,1]`) compared to "Low" (`r x[28,1]`) stiffness.
The arc from clevel $\rightarrow$ balance_1 had a strength of 0.51, with lower level performers more likely to have decreased balance ability (`r x[21,1]`), compared to national level athletes (`r x[22,1]`).
Arcs were also present between independent variables in the network. Strong arcs were present between BIS_1 $\rightarrow$ FFFS_1 (0.98) and BIS_2 $\rightarrow$ FFFS_2 (0.68).
In both instances, "High" FFFS was more likely when BIS was "High" (`r y[2,1]` for _1, `r y[6,1]` for _2) compared to "Low" (`r y[1,1]` for _1, `r y[5,1]` for _2).
The arc between nlelg $\rightarrow$  BIS had a strength of 0.62 for nlelg_1  $\rightarrow$ BIS_1, however no arc was present between nlelg_2 and BIS_2.
For nlelg_1 $\rightarrow$ BIS_1, "Low" negative life events increased the probability of BIS being in the "High" state from `r y[13,1]` to `r y[14,1]`.

@@

The Markov blanket for injured\_1, which contained hours spent training per week (hours), negative life events (nlelg_1), muscle stiffness (stiffness_1), current competitive level (clevel) and balance (balance_1), is shown in Fig \ref{fig:fig7}.
The arc between nlelg_1 and injured_1 was fixed in the network, so has the maximum strength of 1.

```{r}
# plotsg1 <- model2network('[hours][stiffness_1][nlelg_1][clevel][injured_1|hours:nlelg_1:stiffness_1][balance_1|clevel:injured_1]')
# q1p <- graphviz.plot(plotsg1,
#                      shape = "rectangle",
#                      render = F,
#                      groups = list(c("clevel", "hours"), "injured_1", c("stiffness_1", "nlelg_1", "balance_1")))
# 
# strq1 <- str %>%
#   filter(from %in% c("clevel", "hours", "nlelg_1", "stiffness_1", "injured_1") &
#            to %in% c("balance_1", "injured_1")) %>%
#   filter(strength > 0.30) %>%
#   mutate(strength = format(round(strength, digits = 2), nsmall = 2))
# 
# labels <- as.character(strq1$strength)
# labels[4] <- str_pad(labels[4], side = "left", width = 2)
# 
# edgenames <- edgeNames(q1p)
# names(labels) <- edgenames
# nodeRenderInfo(q1p)$fill[c("injured_1", "nlelg_1")] = "gold"
# edgeRenderInfo(q1p) <- list(lwd = c("nlelg_1~injured_1" = 6,
#                                     "stiffness_1~injured_1" = 3,
#                                     "clevel~balance_1" = 3),
#                             lty = c("injured_1~balance_1" = "dashed",
#                                     "hours~injured_1" = "dashed"))
# 
# q1p <- layoutGraph(q1p, edgeAttrs=list(label = labels))
# graph.par(list(edges=list(fontsize = 12)))
# 
# nodeRenderInfo(q1p)$fill[c("nlelg_1", "stiffness_1", "balance_1")] = "deepskyblue"
# nodeRenderInfo(q1p)$fill["injured_1"] = "gold"
# renderGraph(q1p)

The conditional probability query (CPQ) for injured_1 in the "injured" state for all variables that were directly linked to injured_1 is shown in Table \ref{tab:table5}.
The probability of injured_1 = "injured" rose from `r query1[4,2]` to `r query1[4,3]` when stiffness was "High" compared to "Low".
Negative life events had a negligible effect when moving from the "Low" to "High" state.

@@

The second CPQ investigated the outcome of injured_1 being "injured" conditional on all variables in the Markov blanket.
The Markov blanket contained five nodes, each with two possible states resulting in $2^5$ combinations of variables, therefore only the three lowest and highest probabilities are shown in Table \ref{tab:table6} (complete results in S4 Table).
The combination of lower competitive level, "High" hours per week, "Low" negative life events, "High" balance and "High" stiffness resulted in a probability of `r max(query2$prob)` for injured_1 being in the "injured" state.
When all variables were in the "Low" state the probability of "injured" was approximately `r min(query2$prob)`
Negative life events only had a substantial effect on injured_1 when all other variable were fixed to "Low".
In this instance the probability of injured_1 being "injured" rose from 0.04 to 0.19, when negative life events was in the "Low" and "High" states respectively.

@@

query2 %>%
  select(7, 2:6) %>%
  rename(Probability = "prob") %>%
  `colnames<-`(str_replace_all(colnames(.), "_", "\\\\_")) %>%
  mutate(clevel = str_replace_all(clevel, "_", "\\\\_")) %>%
  kable(
    .,
    caption = "Highest and lowest probability of injured\\_1 being in the 'injured' state, conditional on the all variables in the Markov blanket for injured\\_1.",
    escape = F
  ) %>%
  kable_styling(latex_options = "HOLD_position") %>%
  pack_rows("Highest", 1, 3) %>%
  pack_rows("Lowest", 4, 6) %>%
  row_spec(0,bold=TRUE)
```

<!-- Table \ref{tab:table7} shows the states of all explanatory and independent variables suffixed with _1 that resulted in the highest and lowest probability of injured_1 being in the "injured" state. -->
<!-- The probability of injured_1 being "injured" rose from `r injured1comp[2,3]` to `r injured1comp[2,2]` with the combination of variable states in Table \ref{tab:table7}.  -->
<!-- Notably, hours per week, stiffness_1 and balance_1 had the greatest effect on raising the probability of injured_1 being in the "injured" state. -->

>>>>

floatsintext      : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
csl: frontiers.csl
always_allow_html: true
header_includes:
  - \usepackage{pdflscape}
  - \newcommand{\blandscape}{\begin{landscape}}
  - \newcommand{\elandscape}{\end{landscape}}
---

@@

```{r include=FALSE}
options(scipen=999, digits = 2)
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
 source("r-files/functions.R")
 source("r-files/methods_code.R")
 source("r-files/results_code.R")
#load("all_data.RData", envir = myEnv)
#myEnv <- new.env()
#attach(myEnv, name="sourced_scripts")

pacman::p_load(tidyverse, 
               knitr,
               kableExtra,
               qwraps2, 
               brms,
               broom.mixed, 
               bnlearn, 
               tidyselect, 
               Rgraphviz, 
               readxl, 
               janitor,
               irr, 
               lavaan,
               caret,
               parallel,
               zoo, 
               sjmisc)

@@

10422 words in text body
3142 words in reference section

# Introduction

@@

```{r fig1, fig.cap="Stress and injury model (Williams and Andersen, 1998).", fig.align='center',fig.pos = "!h"}
include_graphics("figures_doc/Fig1.jpeg")
```

@@

# Material and Methods

@@

## Ethics Statement

Ethical approval was obtained from the University ethics committee prior to the start of the study and all participants provided informed consent.

## Participants

A total of 351 athletes (male: _n_ = 231, female: _n_ = 120) were initially recruited for the study from a British University and local sports clubs (Table \ref{tab:table1}). Participants had an average age of `r mean_sd(parchar$age, digits = 1)` years and represented a range of team (football, rugby, netball, cricket, lacrosse, basketball and field hockey) and individual sports (athletics, tennis, weightlifting, gymnastics, judo, swimming and golf. 
Participants self-rated competitive level ranged from recreational to international standard. 
They were engaged in training for their respective sports for at least five hours per week.
A total of `r n_perc(parchar$pic == "1")` participants had sustained an injury in the 12 months prior to the start of the study (male: _n_ = 114 [49%], female; _n_ = 48 [40%]).
All participants were injury free (no modifications to their usual training routine due to a sport related medical problem for a minimum of four weeks) at the start of the study.

@@

```{r fig2, fig.cap="mBESS positions (A-F). Top row, firm surface. Bottom row, foam surface. Left column, dominant leg stance. Middle column, non-dominant leg stance. Right column, Tandem leg stance.", out.width = "100%", fig.align='center',fig.pos = "!h"}
include_graphics("figures_doc/Fig2.jpeg")
```

@@

```{r fig3, fig.cap = 'Study design. For each time point (T), each box contains the number of participants who completed the data collection (\\textit{n}), the measures used for data collection and the approximate date of the data collection.', out.width="100%"}

include_graphics("figures_doc/studydesign_new.png")
```

@@

```{r fig4, fig.cap = 'Outline of the protocol for each data collection.', out.width="100%", fig.align='center'}
include_graphics("figures_doc/sessionprotocol.png")
```

@@

```{r fig5, fig.cap="A simple discrete Bayesian network contain nodes, possible states of the nodes and the arcs connecting nodes.", fig.pos="H", out.width="100%", fig.align='center'}

include_graphics("figures_doc/examplenetwork.png")

@@

BNs have recently been used in the sport psychology literature [@Olmedilla2018; @Fuster-Parra2017; @Ruiz-Perez2019] and offer several benefits over traditional statistical analysis. 
For example, predictions can be made about any variable in the network, rather than there being a distinction between dependent and independent variables in the data, such as in linear regression models that are often used within the sport psychology literature [@Olmedilla2018; @Bittencourt2016].
Furthermore, the structure of a network can be obtained from both empirical data _and_ prior knowledge about the area of study; the latter being particularly useful when there are a large number of variables in the network, or only a small number of observations are available in the data [@Xiao-xuan2007].
In such instances, a purely data driven approach to learning the network would be time-consuming due to the large parameter space, and inefficiency at identifying an approximation of the true network structure. 
Prior knowledge about dependencies between variables can therefore be included in the network structure, while still allowing a data driven approach for unknown dependencies, to improve the overall computation of the network structure [@Heckerman1995; @Xu2015].
The following sections detail the steps taken in the current study to firstly prepare the data for each network, and then obtain the structure of each network that was used for further inference.

@@

Of the 351 participants that were initially recruited for the study, 94 only completed the first time point, and therefore had to be removed from the study as no injury information was obtained for these participants following the first time point. 
To prepare the data for the BN, missing values in the dataset were first imputed. 
Out of the 650 total measurements across all time points in the current study, there were `r n_perc(missingtot$stiffness == 0)` missing muscle stiffness measurements and `r n_perc(missingtot$hrv == 0)` missing heart rate recordings. The missing data were due to technical faults in the data collection equipment and were considered to be missing completely at random. 
A missing rate of 15-20% has been reported to be common in psychological studies, and several techniques are available to handle missing values [@Enders2003; @Lang2014].
In the current study, the _caret_ package [@Kuhn2008] was used to impute the missing values. 
A bagged tree model using all of the non-missing data was first generated and then used to predict each missing value in the dataset. 
The bagged tree method is a reliable and accurate method for imputing missing values in data and is superior to other commonly used methods such a median imputation [@Kuhn2008].

A median split technique was used to discretise the data used in the network into "Low" and "High" levels.
All variables apart from negative and total life events were approximately normally distributed and required no further transformation prior to the median split. 
For the LESCA questionnaire data, a cumulative total of the current, and previous time points was calculated at each time point to account for the potential continuing effect of the life events experienced by athletes over time.
Given the limited support for a relationship between positive life events and injury [@Williams2007], only negative and total life events were included in the network.
Cumulative negative, and cumulative total life event scores at each time point were first log scaled so distributions were approximately normal, and then binarised using the median at each time point (nlelg and tlelg respectively).
In addition to the log scaled cumulative values, an untransformed NLE score from the first time point was included as an additional variable based on previous literature that indicates this variable should have a strong relationship with injury outcome [@Ivarsson2017].

@@

To obtain the network structure, several steps were taken to ensure that both a theoretically realistic network, and a network that was an appropriate fit to the collected data, was used for inference. 
Prior knowledge about the network structure was included by providing a list of arcs that are always _restricted_ from being in the network (blacklist), and a list of arcs that are always _included_ in the network (whitelist). 
Additionally, there are several scoring functions such as Bayesian Information Criteria (BIC) and Bayesian Dirichlet equivalent uniform (BDeu) that can be used to compare network structures with certain nodes and arcs included or excluded [@Scutari2014]. 
To account for the repeated measures design employed in this study and to maximise the use of the data, pairs of complete cases (e.g., participants who completed T1 + T2, and T2 + T3) were used in a two-time Bayesian network (2TBN) structure [@Murphy2002].
In the 2TBN, variables measured at T2 could depend on variables measured at T1 (e.g., T1 $\rightarrow$ T2) and variables measured at T3 could depend on variables measured at T2 (e.g., T2 $\rightarrow$ T3).
However, arcs were blacklisted between T2 $\rightarrow$ T1 and T3 $\rightarrow$ T2 to preserve the order in which data was collected. 
Variables were separated into two groups; "explanatory", for variables that were fixed (e.g., gender), or "independent", for variables that were measured at each time point and could vary during the study.
Independent variable names were suffixed with _1 for time point T, and _2 for time point T+1 (e.g., T1_1 $\rightarrow$ T2_2 and T2_1 $\rightarrow$ T3_2).
Formatting the data in this way meant participants who completed T1 and T2, but did not complete T3, could still be included in the analysis. 
Table \ref{tab:table2} provides an example of the formatted data and demonstrates that participants 1 and 3 have complete data, and therefore have two rows of data each representing variables from T1 $\rightarrow$ T2 and T2 $\rightarrow$ T3, respectively. 
Participant 2 did not complete the final data collection at T3 and therefore only has one row of data representing the variables collected at T1 and T2.
In addition to the blacklisted arcs between T2 $\rightarrow$ T1 and T3 $\rightarrow$ T2, the direction of arcs was restricted between independent variables and explanatory variables (e.g., independent $\rightarrow$ explanatory); however, arcs were not restricted between explanatory $\rightarrow$ independent variables.
Finally, arc direction was restricted between specific nodes within the explanatory variables. 
Arcs from clevel $\rightarrow$ gender, nlebase $\rightarrow$ gender and nlebase $\rightarrow$ ind_team were included in the blacklist, as arcs in these directions did not make logical sense. 
All subsequent models used the same blacklist.

@@

The subscales representing the BAS (RR, RI, GDP and I) showed limited connection to other variables in the network. 
Therefore, competing models were examined and BIC scores compared to establish the model with the best fit to the data (values are shown relative to the highest value).
RI provided the highest BIC value (`r rstmodel[1,2]`), compared to RR (`r rstmodel[2,2]- rstmodel[1,2]`), GDP (`r rstmodel[4,2]-  rstmodel[1,2]`) and I (`r rstmodel[5,2] - rstmodel[1,2]`).
Including all the variables (RR, RI, GDP and I) resulted in a significantly lower score `r rstmodel[6,2] - rstmodel[1,2]`) indicating that including all the variables was not beneficial to the model structure and did not offset the cost of the additional parameters. Therefore, only RI was included in the final structure. 
Finally, both total score and asymmetry for balance were included in the initial network. However, visual inspection of the network revealed no arcs between bal_asym_1 or bal_asym_2 and any other node in the network.
Therefore, balance asymmetry was removed from the final network structure.
To summarise, Table \ref{tab:table3} reports the variables that were included in the final network structure.

```{r table3, results='asis'}
varcutoffs %>% 
  filter(!var %in% c("sdnn", "RR", "I", "GDP", "tlelg", "tlelg_1", "tlelg_2", "tlelg_3", "bal_asym")
) %>% 
  mutate(var = ifelse(var == "hrv", "HRV", var)) %>% 
  kable("latex",
       align = c("llcc"),
       caption = "Variables included in the final Bayesian network structure.",
       col.names = c("Variable", "Definition", "State 1", "State 2")) %>%
  kable_styling(latex_options = "HOLD_position") %>% 
  column_spec(2, width = c("5.2cm")) 
```

@@

```{r include=FALSE}
datanet %>% 
  group_by(id) %>% 
  mutate(count = sum(ifelse(injured == "injured", 1, 0))) %>% 
  select(id, count) %>% 
  mutate(count = ifelse(count == 0, 0, 1)) %>%
  distinct(id, .keep_all = TRUE) %>%
  ungroup() %>% 
  janitor::tabyl(count)
```

During the study, 46% (_n_ = 117) of participants reported at least one injury with an average severity of `r mean_sd(datanet$days_missed, digits = 0)`, days (range = `r min(datanet$days_missed[datanet$days_missed > 1])` - `r max(datanet$days_missed)` days).
Both male and female participants reported a greater number of acute compared to chronic injuries (male, acute = `r inj_mechanism[1,3]`, chronic = `r inj_mechanism[2,3]`; female, acute = `r inj_mechanism[1,2]` chronic = `r inj_mechanism[2,2]`), and non-contact injuries were more common than contact injuries (male, non-contact = `r inj_mechanism[4,3]`, contact = `r inj_mechanism[2,3]`; female, non-contact = `r inj_mechanism[4,2]` contact = `r inj_mechanism[3,2]`).
Table \ref{tab:table4} shows the number and percentage of injury types sustained by both male and female participants.

@@

y <- pmap_dfr(.l = list(from = rep(from, 2),
     to = rep(to, 2),
     state = rep(c("High", "Low"), each = length(from))),
         ~newprobtable(tallc, vars = ..1, outcome = ..2, state = ..3, avg30)) %>%
  gather(var, state, -prob, -outcome) %>%
  drop_na(state) %>%
  mutate(prob = format(round(prob, digits = 2), nsmall = 2))

@@

cbind(inj_summary[c(1:3), c(2:4)], inj_summary[c(4:6), c(3:4)]) %>%
  kable(
    .,
    caption = "The number and percentage (\\%) of types of injuries sustained by male and female participants.",
    col.names = c("", "Lower body", "Upper body", "Lower body", "Upper body"),
    escape = F
  ) %>%
  add_header_above(c(" " = 1, "Female" = 2, "Male" = 2), bold = T) %>%
  footnote(general = "Other included bone, skin and brain injuries.") %>%
  kable_styling(latex_options = "HOLD_position") %>%
  row_spec(0, bold = TRUE)

str_finder <- function(from, to) {
  format(round(str[which(str$from == from & str$to == to), 3], digits = 2), nsmall = 2)
}

str_cfinder <- function(from, to) {
  format(
    round(
      strchange[which(strchange$from == from &
        strchange$to == to), 3],
      digits = 2
    ),
    nsmall = 2
  )
}
```

@@

The first network structure obtained from the data (Fig \ref{fig:fig6}) examined the interactions between explanatory variables, independent variables and probability of injury across time points in the study. 
Strong sport-related and gender-based connections between several explanatory and independent variables were demonstrated for individual and team-based sports.
The "ind_team" node had strong arcs to hours (`r str_finder("ind_team", "hours")`) and nlebase (`r str_finder("ind_team", "nlebase")`).
Individual athletes were more likely to have "High" hours per week (`r x[5,1]`) compared to team-based athletes (`r x[6,1]`).
Individual athletes were also more likely to have "High" negative life events in the 12 months preceding the start of the study compared to team-based athletes (individual athletes = `r x[7,1]`, team-based athletes = `r x[8,1]`).
The arc from clevel $\rightarrow$ balance_1 had a strength of `r str_finder("clevel", "balance_1")`, with lower level performers more likely to have decreased balance ability (`r x[21,1]`), compared to national level athletes (`r x[22,1]`). 
High gender-based connections were reported for the arcs from gender  $\rightarrow$ stiffness_1 (`r str_finder("gender", "stiffness_1")`) and gender $\rightarrow$ stiffness_2 (`r str_finder("gender", "stiffness_2")`), with males more likely to have "High" stiffness compared to females (males = `r x[14,1]`, females = `r x[13,1]`).
Irrespective of sport or gender, strong connections were found between explanatory variables. The arc from nlebase $\rightarrow$ RI_1 had a strength of `r str_finder("nlebase", "RI_1")`, and the probability of RI_1 being in the “High” state increased from `r x[2,1]` to `r x[3,1]` when nlebase increased from “Low” to “High”.

The first network demonstrated further strong variable interactions between high stiffness, poor balance and injury probability. 
The arc from pi $\rightarrow$ stiffness_1 was `r str_finder("pi", "stiffness_1")` with athletes who reported an injury in the preceding 12 months being more likely to have “High” (`r x[26,1]`) compared to “Low” (`r x[28,1]`). stiffness.
Strong arcs were present between BIS_1 $\rightarrow$ FFFS_1 (`r str_finder("BIS_1", "FFFS_1")`) and BIS_2 $\rightarrow$ FFFS_2 (`r str_finder("BIS_2", "FFFS_2")`) were identified.
In both instances, "High" FFFS was more likely when BIS was "High" (`r y[2,1]` for _1, `r y[6,1]` for _2) compared to "Low" (`r y[1,1]` for _1, `r y[5,1]` for _2).
The arc between nlelg $\rightarrow$  BIS had a strength of `r str_finder("nlelg_1", "BIS_1")` for nlelg_1  $\rightarrow$ BIS_1, and `r str_finder("nlelg_2", "BIS_2")` nlelg_2 and BIS_2.
"Low" negative life events increased the probability of BIS being in the "High" state from `r y[13,1]` to `r y[14,1]` for nlelg_1  $\rightarrow$ BIS_1, and `r y[17,1]` to `r y[18,1]` for nlelg_2  $\rightarrow$ BIS_2.

\newpage

```{r fig6, fig.align='default', fig.height=8, fig.width=12, fig.cap='The full Bayesian network structure was plotted using the _strength.plot_ function in bnlearn. The strength of each arc is shown graphically by the style of the arc. Thin, dashed arcs indicate the weakest arcs (arc strength below 0.50), whereas thick solid arcs indicate the strongest arcs (arc strength of 1). White nodes in the network indicate the explanatory variables, blue nodes indicate T1_1 and T2_1 variables, and red nodes indicated T2_2 and T3_2 variables. The injured_X nodes have been coloured gold as they are the main nodes of interest within the network.'}
renderGraph(gR)
```

@@

The first conditional probability query (CPQ) investigated the variables in the Markov blanket for injured\_1 (Fig \ref{fig:fig7}), which contained hours spent training per week (hours), negative life events (nlelg_1), muscle stiffness (stiffness_1), current competitive level (clevel) and balance (balance_1). 
The arc between nlelg_1 and injured_1 was fixed in the network, so has the maximum strength of 1.

```{r fig7, fig.cap="Markov blanket of injured\\_1. Arc strengths are included as arc labels.", out.width="100%", fig.align="center"}
include_graphics("figures_doc/Fig7.jpeg")
```

The CPQ for injured_1 in the "injured" state for all variables that were linked to injured_1 is shown in Table \ref{tab:table5}.
The probability of injured_1 = "injured" rose from `r query1[4,2]` to `r query1[4,3]` when stiffness was "High" compared to "Low".
Negative life events had a negligible effect when moving from the "Low" to "High" state.

@@

The second CPQ investigated the outcome of injured_1 being "injured" conditional on all variables in the Markov blanket.
The Markov blanket contained five nodes, each with two possible states resulting in $2^5$ combinations of variables, therefore only the three lowest and highest probabilities are shown in Table \ref{tab:table6} (complete results in S4 Table).
The combination of lower competitive level, "High" hours per week, "Low" negative life events, "High" balance and "High" stiffness resulted in a probability of `r max(query2$prob)` for injured_1 being in the "injured" state.
When all variables were in the "Low" state the probability of "injured" was approximately `r min(query2$prob)`

Negative life event stress had a negligible effect on the probability of injury, only influencing  injured_1 when all other variable were fixed to "Low".
In this instance, the probability of injured_1 being "injured" rose marginally from 0.04 to 0.19, when negative life events was in the "Low" and "High" states, respectively.

@@

query2 %>%
  select(7, 2:6) %>%
  rename(Probability = "prob") %>%
  `colnames<-`(str_replace_all(colnames(.), "_", "\\\\_")) %>%
  mutate(clevel = str_replace_all(clevel, "_", "\\\\_")) %>%
  kable(
    .,
    caption = "Highest and lowest probability of injured\\_1 being in the 'injured' state, conditional on the all variables in the Markov blanket for injured\\_1.",
    escape = F,
    align = "ccccc"
  ) %>%
  kable_styling(latex_options = "HOLD_position") %>%
  pack_rows("Highest", 1, 3) %>%
  pack_rows("Lowest", 4, 6) %>%
  row_spec(0,bold=TRUE)
```
